---
title: "Urban Growth Modeling for Pittsburgh Metropolitan Statistical Area"
subtitle: "Planning Memo to Southwestern Pennsylvania Commission"
author: "[Xinyi Qiu](https://xy16-5.github.io/) and [Gillian Xuezhu Zhao](https://gillianzhaoxz.github.io/web/)"
date: "5/10/2021"
output:   
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
    theme: cerulean
---
```{r setup knitr, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE,include=FALSE)
options(scipen=99)
```
```{r setwd, include=FALSE}
#Xinyi's
#setwd("~/Desktop/Courses/Land Use and Environmental Modeling/Final/Data")
#Gillian's
setwd("G:/UPenn/CPLN675_LandUse_and_EnvironmentalModeling/final/CPLN-675-Final/Data")
getwd()
```
```{r Set up, include=FALSE, purl = TRUE}
library(tidyverse)
library(sf)
library(raster)
library(knitr)
library(formattable)  # I use formattable to replace the kableExtra
# library(kableExtra)
library(tidycensus)
library(tigris)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
library(FNN)
library(QuantPsyc)
library(caret)
library(yardstick)
library(pscl)
library(plotROC) 
library(ggrepel)
library(pROC)
library(grid)
library(gridExtra)
library(viridis)
library(exactextractr)
library(stringr)
library(censusapi)
library(ROCR)
library(igraph)
library(stargazer)

plotTheme <- function(base_size = 10) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 12,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=11),
    axis.title = element_text(size=11),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 12)
  )
}

mapTheme <- function(base_size = 10) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 12,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    strip.text.x = element_text(size = 12))
}

palette2 <- c("#f8edeb", "#83c5be")
palette2b <- c("#f8edeb", "#fd9235")
palette4 <- c("#a1dab4","#41b6c4","#2c7fb8","#253494")
palette5 <- c("#f7f6de","#edeec9","#b8dedc","#83c5be","#42999b")
palette5b <- c("#42999b","#83c5be","#b8dedc","#edeec9","#f7f6de")
palette10 <- c("#f7fcf0","#e0f3db","#ccebc5","#a8ddb5","#7bccc4",
               "#4eb3d3","#2b8cbe","#0868ac","#084081","#f7fcf0")
```
```{r function}
quintileBreaks <- function(df,variable) {
    as.character(quantile(df[[variable]],
                          c(.01,.2,.4,.6,.8),na.rm=T))
}

#This function can be used to convert a polygon sf to centroids xy coords.
xyC <- function(aPolygonSF) {
  as.data.frame(
    cbind(x=st_coordinates(st_centroid(aPolygonSF))[,1],
          y=st_coordinates(st_centroid(aPolygonSF))[,2]))
} 

#this function convert a raster to a data frame so it can be plotted in ggplot
rast <- function(inRaster) {
  data.frame(
    xyFromCell(inRaster, 1:ncell(inRaster)), 
    value = getValues(inRaster)) }

nn_function <- function(measureFrom,measureTo,k) {
  #convert the sf layers to matrices
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
    output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}

aggregateRaster <- function(inputRasterList, theFishnet) {
  #create an empty fishnet with the same dimensions as the input fishnet
  theseFishnets <- theFishnet %>% dplyr::select()
  #for each raster in the raster list
  for (i in inputRasterList) {
  #create a variable name corresponding to the ith raster
  varName <- names(i)
  #convert raster to points as an sf
    thesePoints <-
      rasterToPoints(i) %>%
      as.data.frame() %>%
      st_as_sf(coords = c("x", "y"), crs = st_crs(theFishnet)) %>%
      filter(.[[1]] == 1)
  #aggregate to the fishnet
    thisFishnet <-
      aggregate(thesePoints, theFishnet, length) %>%
      mutate(!!varName := ifelse(is.na(.[[1]]),0,1))
  #add to the larger fishnet
    theseFishnets <- cbind(theseFishnets,thisFishnet)
  }
  #output all aggregates as one large fishnet
   return(theseFishnets)
  }
```
```{r load pittsburgh boundary, include=FALSE}
Pittsburgh_MSA <-
  st_read("Pittsburgh_MSA/Pittsburgh_MSA.shp") %>%
  st_transform("ESRI:102741") #102003

Pittsburgh_MSA_boundary <- Pittsburgh_MSA %>%
  group_by(STATEFP) %>%
  summarise() %>%
  st_transform("ESRI:102741")

MSA_fishnet <- 
  st_make_grid(Pittsburgh_MSA_boundary, 3000) %>%
  st_sf()

MSA_fishnet <- 
  MSA_fishnet[Pittsburgh_MSA_boundary,]
```

# Introduction
Since its days as the Forge of America, Pittsburgh, the biggest city in Pittsburgh MSA, has changed dramatically in its over 250-year history. With decreasing population, decayed structures, and yet suburban sprawl, the region is eager to redevelop many of its areas and neighborhoods^[Lamorella, J. (2013). Crisis + opportunity. Re*imagining Pittsburgh. http://pittsburghcriticalpoints.weebly.com/crisis.html].

We completed the growth modeling project in anticipation of the Southwestern Pennsylvania Commission (SPC)’s next large-scale comprehensive planning process for Pittsburgh Metropolitan Statistical Area (MSA). 

```{r county plot, include=T}
ggplot()+
  labs(title="Pittsburgh MSA Boundary and Counties",
       subtitle="Illustration of Fishnet, 3000 foot resolution")+
  geom_sf(data=Pittsburgh_MSA, aes(fill=as.factor(NAME)))+
  geom_sf(data=MSA_fishnet, fill=NA, color="gray", alpha=0.5) +
  geom_sf(data=Pittsburgh_MSA, fill=NA, size=1.5, aes(colour=as.factor(NAME)))+
  scale_color_viridis(discrete = T,
                     name ="County")+
  scale_fill_viridis(discrete = T,
                     name ="County")+
  mapTheme()
```

We forecasted change in development for 2021 with patterns from 2001-2011 through a binary logistic regression model per each 3000 foot fishnet grid. 

```{r load lc data}
lc_2001 <- raster("msa_2001lc.tif")
lc_2001 <-projectRaster(lc_2001,crs=crs(Pittsburgh_MSA))

lc_2011 <- raster ("msa_2011lc.tif")
lc_2011 <- projectRaster(lc_2011,crs=crs(Pittsburgh_MSA) )

# aggregate
lc_2001 <- raster::aggregate(lc_2001, fact = 30, fun = modal)
lc_2011 <- raster::aggregate(lc_2011, fact = 30, fun = modal)

# calculate
lc_change <- (lc_2001 != lc_2011) * lc_2011
```
```{r lc plot}
reclassMatrix <- 
  matrix(c(
    0,12,0,
    12,24,1,
    24,Inf,0),
    ncol=3, byrow=T)
lc_change_reclass <- reclassify(lc_change, reclassMatrix)
lc_change_reclass[lc_change_reclass < 1] <- 0
names(lc_change_reclass) <- "lc_change_reclass"

lc_change_extract <- data.frame(exact_extract(lc_change_reclass, MSA_fishnet, fun = "mode"))
names(lc_change_extract) <- "developed_0111"

MSA_fishnet01 <- cbind(data.frame(MSA_fishnet),lc_change_extract)

MSA_fishnet01 <- 
  MSA_fishnet01 %>%
  st_as_sf(.) %>%
  st_transform("ESRI:102741")
```

[] Land Use change plot

```{r dev plot, include=T}
ggplot() +
  geom_sf(data=MSA_fishnet01,aes(fill=as.factor(developed_0111)),color='transparent')+
  scale_fill_manual(values=palette2b, name ="",labels=c("No Change","New Development"))+
  labs(title="Pittsburgh MSA Development Land Use Change (2001 - 2011)") +
  mapTheme()
```

[] Land cover types plot

```{r lc type, include=T}
ggplot() +
  geom_sf(data=Pittsburgh_MSA_boundary) +
  geom_raster(data=rast(lc_2001) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(round(value,0)))) +
  scale_fill_viridis(discrete=TRUE,direction=-1, name ="") +
  labs(title = "Pittsburgh MSA Land Cover, 2001",
       subtitle = "Land cover value dictionary see appendix") +
  mapTheme()
```

```{r lc reclass}
developed <- lc_2001 == 21 | lc_2001 == 22 | lc_2001 == 23 | lc_2001 == 24
forest <- lc_2001 == 41 | lc_2001 == 42 | lc_2001 == 43 
farm <- lc_2001 == 81 | lc_2001 == 82 
wetlands <- lc_2001 == 90 | lc_2001 == 95 
otherUndeveloped <- lc_2001 == 52 | lc_2001 == 71 | lc_2001 == 31 
water <- lc_2001 == 11
names(developed) <- "developed"
names(forest) <- "forest"
names(farm) <- "farm"
names(wetlands) <- "wetlands"
names(otherUndeveloped) <- "otherUndeveloped"
names(water) <- "water"

layer_list <- list(developed, wetlands, forest, farm, otherUndeveloped, water)
names(layer_list) <- c("developed_2001", "wetlands_2001", "forest_2001", 
                       "farm_2001", "otherUndeveloped_2001", "water_2001")

fish_extract <- function(fishnet, layers) {
  extract_list <- exact_extract(layers, fishnet, fun = "mode")
  return(extract_list)
}


lc_2001_extracts <- lapply(layer_list, fish_extract, fishnet = MSA_fishnet01)
lc_2001_extracts <- data.frame(do.call(cbind, args = lc_2001_extracts))
MSA_fishnet01 <- cbind(data.frame(MSA_fishnet01), lc_2001_extracts)

MSA_fishnet <- 
  MSA_fishnet %>%
  st_as_sf(.) %>%
  st_transform("ESRI:102741")

lc_2001_lcvar <-
  MSA_fishnet01 %>%
  gather(var,value,developed_2001:water_2001) %>%
  dplyr::select(var,value,geometry) %>%
  st_as_sf()

lc_2001_lcvar$var <- as.factor(str_sub(lc_2001_lcvar$var,1,-6))
```
```{r lc type plot1, include=T}
ggplot()+
  geom_sf(data=lc_2001_lcvar,aes(fill=as.factor(value)),color='transparent')+
  facet_wrap(~var)+
  scale_fill_manual(values=palette2, labels=c("Other","Land Cover"),name ="")+
  labs(title = "Land cover types, 2001",
       subtitle = "As fishnet centroids") +
  theme(legend.position = "bottom")+
  mapTheme()
```

[] Development change overlay with

```{r census}
MSAPop10 <-
  get_decennial(geography = "tract",variables = "P001001",year=2010,
                state=42,county=c("Allegheny County","Armstrong County","Beaver County","Butler County","Fayette County",
                                  "Washington County","Westmoreland County"),geometry=TRUE) %>%
  rename(pop_2010 = value) %>%
  st_transform(st_crs(MSA_fishnet))

Sys.setenv(CENSUS_KEY = "2c3e9f9d2d65abab5f7b81fe418054415d363a43")
MSAPop00 <-getCensus(name = "dec/sf1", 
                  vintage = 2000, 
                  vars = c("NAME","P001001"),
                  region = "tract:*", 
                  regionin = "state:42+county:003,005,007,019,051,125,129")
MSAPop00$label <- paste0(MSAPop00$county,"0",MSAPop00$tract)
for (i in 1:721){
  if(nchar(MSAPop00[i,]$label)!=10){
    MSAPop00[i,]$label <-paste0(MSAPop00[i,]$label,"00")
  }
  
}

MSAtracts00 <- st_read("MSAtracts/MSAtracts.shp") %>%
  st_transform("ESRI:102741")

MSAtracts00$tract <-substr(MSAtracts00$GISJOIN2 ,4,13)
MSAtracts00$tract <- as.character(MSAtracts00$tract)

MSAPop00 <- left_join(MSAtracts00, MSAPop00,by=c("tract"="label"))
MSAPop00 <- MSAPop00 %>%
  dplyr::select(GISJOIN2,NAME,P001001,geometry) %>%
  rename(GEOID = GISJOIN2,pop_2000 = P001001)
```

  [] Population decline plot (optional) and comparison map from Census data
  
```{r pop0010 plot, include=T}
grid.arrange(
ggplot() +
  geom_sf(data = MSAPop00, aes(fill=factor(ntile(pop_2000,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=quintileBreaks(MSAPop00,"pop_2000"),
                   name="Quintile\nBreaks") +
  labs(title="Population, Pittsburgh MSA: 2000") +
  mapTheme(),

ggplot() +
  geom_sf(data = MSAPop10, aes(fill=factor(ntile(pop_2010,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=quintileBreaks(MSAPop10,"pop_2010"),
                   name="Quintile\nBreaks") +
  labs(title="Population, Pittsburgh MSA: 2010") +
  mapTheme(), ncol=2)
```


```{r pop0010 fishnet}
MSA_fishnet01 <-
  MSA_fishnet01 %>%
  rownames_to_column("fishnetID") %>%
  mutate(fishnetID = as.numeric(fishnetID)) 

fishnetPop00 <-
  st_interpolate_aw(MSAPop00["pop_2000"],MSA_fishnet01,extensive=TRUE) %>%
  as.data.frame(.) %>%
  left_join(MSA_fishnet01,.,by="geometry") %>%
  mutate(pop_2000=replace_na(pop_2000,0)) %>%
  dplyr::select(fishnetID,pop_2000)

fishnetPop10 <-
  st_interpolate_aw(MSAPop10["pop_2010"],MSA_fishnet01,extensive=TRUE) %>%
  as.data.frame(.) %>%
  left_join(MSA_fishnet01,.,by="geometry") %>%
  mutate(pop_2010=replace_na(pop_2010,0)) %>%
  dplyr::select(pop_2010,fishnetID)

fishnetPop <- 
  left_join(fishnetPop00,fishnetPop10, by="fishnetID") %>%
  dplyr::select(pop_2000,pop_2010,fishnetID) %>%
  mutate(pop_Change = pop_2010 - pop_2000)


MSA_fishnet01 <- left_join(MSA_fishnet01,data.frame(fishnetPop),by="fishnetID") %>%
  st_as_sf()
```

```{r pop0010 fishnet plot, include=T}
grid.arrange(
ggplot() +
  geom_sf(data=MSAPop10, aes(fill=factor(ntile(pop_2010,5))),colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=substr(quintileBreaks(MSAPop10,"pop_2010"),1,4),
                   name="Quintile\nBreaks") +
  labs(title="Population, Pittsburgh MSA: 2010",
       subtitle="Represented as tracts; Boundaries omitted") +
  mapTheme(),

ggplot() +
  geom_sf(data=MSA_fishnet01, aes(fill=factor(ntile(pop_2010,5))),colour=NA) +
  scale_fill_manual(values = palette5,
                   labels=substr(quintileBreaks(fishnetPop,"pop_2010"),1,4),
                   name="Quintile\nBreaks") +
  labs(title="Population, Pittsburgh MSA: 2010",
       subtitle="Represented as fishnet gridcells; Boundaries omitted") +
  mapTheme(), ncol=2)
```

```{r popchange fishnet plot, include=T}
ggplot() +
  geom_sf(data = MSA_fishnet01, aes(fill=factor(ntile(pop_Change,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=substr(quintileBreaks(MSA_fishnet01,"pop_Change"), 0,4),
                   name="Quintile\nBreaks") +
  labs(title="Population change, Pittsburgh MSA: 2000-2010") +
  mapTheme()
```


  [] Dist to highway
  
```{r highway1, include=FALSE}
PittsburghHighways <-
  st_read("MSAHighways/MSAHighways.shp") %>%
  st_transform(st_crs(Pittsburgh_MSA)) %>%
  st_intersection(Pittsburgh_MSA_boundary)

PittsburghHighways <- 
  PittsburghHighways %>%
  dplyr::select(CTY_CODE, DISTRICT_N,MAINT_FUNC,geometry)
```
```{r highway plot, include=T}
palette2new <- c("NA", "#fd9235")

ggplot() +
  geom_sf(data=Pittsburgh_MSA, color=NA, fill="#f8edeb")+
  geom_sf(data=PittsburghHighways, color='#6c757d', size=1, alpha=0.5) +
  geom_sf(data=MSA_fishnet01, aes(fill=as.factor(developed_0111)),color='transparent') +
  scale_fill_manual(values=palette2new, name ="Land Cover\nChange",labels=c("No Change","New Development")) +
  labs(title = "Pittsburgh MSA New Development and Highways") +
  mapTheme()
```

Our findings aim to help planners model the interplay between supply and demand-side insights and guide new development across to places where growth can have economic and sustainable impacts.

# Exploratory Analysis
## Feature Engineering

[] Distance to highway

```{r dist to highway}
emptyRaster <- lc_change
emptyRaster[] <- NA

highway_raster <- 
  as(PittsburghHighways,'Spatial') %>%
  rasterize(.,emptyRaster)

# writeRaster(highway_raster,"highway_raster.tif")

highway_raster_distance <- distance(highway_raster)
names(highway_raster_distance) <- "distance_highways"

highwayPoints <-
  rasterToPoints(highway_raster_distance) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(MSA_fishnet01))

highwayPoints_fishnet <- 
  aggregate(highwayPoints, MSA_fishnet01, mean) %>%
  mutate(distance_highways = ifelse(is.na(distance_highways),0,distance_highways)) %>%
  rownames_to_column("fishnetID") %>%
  mutate(fishnetID = as.numeric(fishnetID)) 
  

highwayPoints_fishnet <-highwayPoints_fishnet %>%
  st_drop_geometry(.)

MSA_fishnet01 <- left_join(MSA_fishnet01,highwayPoints_fishnet,by="fishnetID") %>%
  st_as_sf()
```
```{r dist to highway plot, include=T}
ggplot() +
  geom_sf(data=Pittsburgh_MSA_boundary) +
  geom_point(data=MSA_fishnet01 , aes(x=xyC(MSA_fishnet01 )[,1], 
                                             y=xyC(MSA_fishnet01)[,2], 
                 colour=factor(ntile(distance_highways,5))),size=1.5) +
  scale_colour_manual(values = palette5,
                      labels=substr(quintileBreaks(MSA_fishnet01 ,"distance_highways"),1,8),
                      name="Quintile\nBreaks") +
  geom_sf(data=PittsburghHighways, color='#6c757d', size=1, alpha=0.8) +
  labs(title = "Distance to Highways",
       subtitle = "As fishnet centroids; Highways visualized in red") +
  mapTheme()
```

[] Spatial lag of dev

```{r spatial lag dev}
MSA_fishnet01$lagDevelopment <-
  nn_function(xyC(MSA_fishnet01),
              xyC(filter(MSA_fishnet01,developed_2001==1)),
              2)
```
```{r spatial lag dev plot, include=T}
ggplot() +
  geom_sf(data=Pittsburgh_MSA_boundary) +
  geom_point(data=MSA_fishnet01, 
             aes(x=xyC(MSA_fishnet01)[,1], y=xyC(MSA_fishnet01)[,2], 
                 colour=factor(ntile(lagDevelopment,5))), size=1.5) +
  scale_colour_manual(values = palette5,
                     labels=substr(quintileBreaks(MSA_fishnet01, "lagDevelopment"),1,7),
                     name="Quintile\nBreaks") +
  labs(title = "Spatial lag to 2001 development",
       subtitle = "As fishnet centroids") +
  mapTheme() +
  geom_sf(data=MSA_fishnet01, aes(fill=as.factor(developed_0111)),color='transparent', alpha=0.8) +
  scale_fill_manual(values=palette2new, name ="Land Cover\nChange",labels=c("No Change","New Development")) 
```
```{r final dataset}
colnames(MSA_fishnet01)[which (names(MSA_fishnet01)=="developed_0111")] <-"lc_change"
```
```{r dataset fishnet}
if ( any("pop_2000.x" == names(MSA_fishnet01))) {
  MSA_fishnet01 <- 
    MSA_fishnet01 %>%
    mutate(pop_2000 = pop_2000.x,
         pop_2010 = pop_2010.x,
         pop_Change = pop_Change.x)
} # this happens sometimes

dat <-
  MSA_fishnet01 %>%
  dplyr::select(lc_change, developed_2001,wetlands_2001,forest_2001,farm_2001,otherUndeveloped_2001,water_2001,pop_2000,pop_2010, pop_Change,
                distance_highways, lagDevelopment)  %>%
  st_join(Pittsburgh_MSA) %>%
  mutate(developed11 = ifelse(lc_change==1 & developed_2001 == 1, 0, developed_2001)) %>%
  filter(water_2001 == 0)

dat$lc_change <- as.factor(dat$lc_change)
```

[] All existing comparison plots

```{r, include=T}
dat %>%
  dplyr::select(distance_highways,lagDevelopment,lc_change) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun = "mean") +
    facet_wrap(~Variable) +
    scale_fill_manual(values = palette2b,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New development as a function of the countinuous variables") +
    plotTheme() 
```

```{r, include=T}
dat %>%
  dplyr::select(pop_2000,pop_2010,pop_Change,lc_change) %>%
  mutate(pop_Change=pop_Change) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun = "mean") +
    facet_wrap(~Variable, scales="free") +
    scale_fill_manual(values = palette2b,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New development as a function of factor variables",
         subtitle = "Original Population Change") +
    plotTheme()
```

```{r, include=T}
table <-dat %>%
  dplyr::select(lc_change:otherUndeveloped_2001) %>%
  gather(Land_Cover_Type, Value, -lc_change, -geometry) %>%
   st_set_geometry(NULL) %>%
     group_by(lc_change, Land_Cover_Type) %>%
     summarize(n = sum(as.numeric(Value))) %>%
     ungroup() %>%
    mutate(Conversion_Rate = paste0(round(100 * n/sum(n), 2), "%")) %>%
    filter(lc_change == 1) %>%
  dplyr::select(Land_Cover_Type,Conversion_Rate) 
table <- data.frame(table)
table$Land_Cover_Type[1]<-"Developed"
table$Land_Cover_Type[2]<-"Farm"
table$Land_Cover_Type[3]<-"Forest"
table$Land_Cover_Type[4]<-"Other Undeveloped"
formattable(table,
            align=c("l","l"))
```

# Modeling
## Model Selection
[] six models

```{r model}
addTaskCallback(function(...) {set.seed(2021);TRUE})

trainIndex <- 
  createDataPartition(dat$developed_2001, p = .50,
                                  list = FALSE,
                                  times = 1)
datTrain <- dat[ trainIndex,]
datTest  <- dat[-trainIndex,]

Model1 <- glm(lc_change ~ wetlands_2001 + forest_2001 + farm_2001 + otherUndeveloped_2001, 
              family="binomial"(link="logit"), data = datTrain)

Model2 <- glm(lc_change ~ wetlands_2001 + forest_2001  + farm_2001 + otherUndeveloped_2001 + lagDevelopment, 
              family="binomial"(link="logit"), data = datTrain)
              
Model3 <- glm(lc_change ~ wetlands_2001 + forest_2001  + farm_2001 + otherUndeveloped_2001 + lagDevelopment + pop_2000, 
              family="binomial"(link="logit"), data = datTrain)          
              
Model4 <- glm(lc_change ~ wetlands_2001 + forest_2001  + farm_2001 + otherUndeveloped_2001 + lagDevelopment + pop_2000 + 
              pop_2010, 
              family="binomial"(link="logit"), data = datTrain)              
            
Model5 <- glm(lc_change ~ wetlands_2001 + forest_2001  + farm_2001 + otherUndeveloped_2001 + lagDevelopment + pop_Change, 
              family="binomial"(link="logit"), data = datTrain)              
              
Model6 <- glm(lc_change ~ wetlands_2001 + forest_2001  + farm_2001 + otherUndeveloped_2001 + lagDevelopment + pop_Change + 
              distance_highways, 
              family="binomial"(link="logit"), data = datTrain) 
```

```{r stargazer, results='asis', include=T}
stargazer(Model1, Model2, Model3, Model4, Model5, Model6, type='html')
```

[] R squared plot

```{r prep r2}
modelList <- paste0("Model", 1:6)
models <- map_dfc(modelList, function(x)pR2(get(x)))[4,] %>%
  setNames(paste0("Model",1:6)) %>%
  gather(Model,McFadden) %>%
  as.data.frame()
```

```{r compare models, include=T}
ggplot(models, aes(Model, McFadden)) +
  geom_bar(stat="identity", fill="#b8dedc") +
  labs(title= "McFadden R-Squared by Model") +
  plotTheme()
```

## Model Outcome {.tabset .tabset-fade .tabset-pills}

[] Probability histogram

```{r histogram prob}
testSetProbs <- 
  data.frame(class = datTest$lc_change,
             probs = predict(Model6, datTest, type="response")) 
```

### Original plot
```{r density outcome, include=T}
ggplot(testSetProbs, aes(probs)) +
  geom_density(aes(fill=class), alpha=0.5) +
  scale_fill_manual(values = palette2,
                    labels=c("No Change","New Development")) +
  labs(title = "Density plot of test set predicted probabilities",
       x="Predicted Probabilities",y="Density") +
  plotTheme()
```

### Zoomed-in plot
```{r density outcome zoomed, include=T}
# zoom closer
ggplot(testSetProbs, aes(probs)) +
  geom_density(aes(fill=class), alpha=0.5) +
  xlim(0, 0.13)+
  ylim(0, 100)+
  scale_fill_manual(values = palette2,
                    labels=c("No Change","New Development")) +
  labs(title = "Density plot of test set predicted probabilities",
       x="Predicted Probabilities",y="Density") +
  plotTheme()
```

## {-}

[] Probability threshold selection table

```{r prob threshold selection}
options(yardstick.event_first = FALSE)

classProbs <- predict (Model6, datTest, type='response')

testProbs <- data.frame (obs = as.numeric(datTest$developed11), pred = classProbs)

pred <- prediction(testProbs[is.na(testProbs$pred)==FALSE,]$pred,testProbs[is.na(testProbs$pred)==FALSE,]$obs)
f.perf<-performance(pred,"f")
plot(f.perf)

F.score <-c(f.perf@y.values[[1]])
cutoff<-c(f.perf@x.values[[1]])
F.score_table<-data.frame(cbind(F.score,cutoff))
F.score_table[which.max(F.score_table$F.score),] #0.0122;

testSetProbs <- 
  testSetProbs %>% 
  mutate(
        predClass_01 = as.factor(ifelse(testSetProbs$probs >= 0.01 ,1,0)),
        predClass_02 = as.factor(ifelse(testSetProbs$probs >= 0.02 ,1,0)),
        predClass_03 = as.factor(ifelse(testSetProbs$probs >= 0.03 ,1,0)), # V
        predClass_05 = as.factor(ifelse(testSetProbs$probs >= 0.05 ,1,0))
         ) 
```
```{r prob table, include=T}
testSetProbs %>%
  dplyr::select(-probs) %>%
  gather(Variable, Value, -class) %>%
  group_by(Variable) %>%
  summarize(Sensitivity = round(yardstick::sens_vec(class,factor(Value)),2),
            Specificity = round(yardstick::spec_vec(class,factor(Value)),2),
            Accuracy = round(yardstick::accuracy_vec(class,factor(Value)),2)) %>% 
  formattable(table,
            align=c("l","l"))
```
```{r prob map, include=T}
predsForMap <-         
  dat %>%
    mutate(probs = predict(Model6, dat, type="response") ,
           Threshold_1_Pct = as.factor(ifelse(probs >= 0.01 ,1,0)),
           Threshold_3_Pct =  as.factor(ifelse(probs >= 0.03, 1,0))) %>%
    dplyr::select(lc_change,Threshold_1_Pct,Threshold_3_Pct) %>%
    gather(Variable,Value, -geometry) %>%
    st_cast("POLYGON")

ggplot() +
  geom_point(data=predsForMap, aes(x=xyC(predsForMap)[,1], y=xyC(predsForMap)[,2], colour=Value)) +
  facet_wrap(~Variable) +
  scale_colour_manual(values = palette2b, labels=c("No Change","New Development"),
                      name="") +
  labs(title="Development predictions - Threshold comparison") + 
  theme(legend.position="bottom") +
  mapTheme()
```
```{r confusion matrix map, include=T}
ConfusionMatrix.metrics <-
  dat %>%
    mutate(probs = predict(Model6, dat, type="response") ,
           Threshold_1_Pct = as.factor(ifelse(probs >= 0.01 ,1,0)),
           Threshold_3_Pct =  as.factor(ifelse(probs >= 0.03, 1,0))) %>%
    mutate(TrueP_1 = ifelse(lc_change  == 1 & Threshold_1_Pct == 1, 1,0),
           TrueN_1 = ifelse(lc_change  == 0 & Threshold_1_Pct == 0, 1,0),
           TrueP_3 = ifelse(lc_change  == 1 & Threshold_3_Pct == 1, 1,0),
           TrueN_3 = ifelse(lc_change  == 0 & Threshold_3_Pct == 0, 1,0)) %>%
    dplyr::select(., starts_with("True")) %>%
    gather(Variable, Value, -geometry) %>%
    st_cast("POLYGON") 

ggplot(data=ConfusionMatrix.metrics) +
  geom_point(aes(x=xyC(ConfusionMatrix.metrics)[,1], 
                 y=xyC(ConfusionMatrix.metrics)[,2], colour = as.factor(Value))) +
  facet_wrap(~Variable) +
  scale_colour_manual(values = palette2, labels=c("Correct","Incorrect"),
                       name="") +
  labs(title="Development predictions - map by confusion matrix") + 
  theme(legend.position="bottom") +
  mapTheme()
```

## Model Validation

```{r spatial CV}
spatialCV <- function(dataFrame, uniqueID, dependentVariable, modelName) {

#initialize a data frame 
endList <- list()

#create a list that is all the spatial group unqiue ids in the data frame (ie counties)    
  uniqueID_List <- unique(dataFrame[[uniqueID]])  
  x <- 1
  y <- length(uniqueID_List)
  
#create a counter and while it is less than the number of counties...  
  while(x <= y) 
  {
#call a current county    
    currentUniqueID <- uniqueID_List[x]
#create a training set comprised of units not in that county and a test set of units
#that are that county
    training <- dataFrame[ which(dataFrame[[uniqueID]] != uniqueID_List[x]),]
    testing <- dataFrame[ which(dataFrame[[uniqueID]] == uniqueID_List[x]),]
#create seperate xy vectors
    trainingX <- training[ , -which(names(training) %in% c(dependentVariable))]
    testingX <- testing[ , -which(names(testing) %in% c(dependentVariable))]
    
    trainY <- training[[dependentVariable]]
    testY <- testing[[dependentVariable]]
#Calculate predictions on the test county as part of a data frame including the observed
#outcome and the unique county ID    
   thisPrediction <- 
     data.frame(class = testY,
                probs = predict(modelName, testingX, type="response"),
                county = currentUniqueID) 

#Row bind the predictions to a data farme
   endList <- rbind(endList, thisPrediction)
#iterate counter    
    x <- x + 1 
  } 
#return the final list of counties and associated predictions  
  return (as.data.frame(endList))
}
```
```{r spatialCV counties}
spatialCV_counties <-
  spatialCV(dat,"NAME","lc_change", Model6) %>%
  mutate(predClass = as.factor(ifelse(probs >= 0.03 ,1,0)))
```
```{r spatialCV table, include=T}
spatialCV_metrics <-
  spatialCV_counties %>% 
    group_by(county) %>% 
    summarize(Observed_Change = sum(as.numeric(as.character(class))),
              Sensitivity = round(yardstick::sens_vec(class,predClass),2),
              Specificity = round(yardstick::spec_vec(class,predClass),2),
              Accuracy = round(yardstick::accuracy_vec(class,predClass),2)) 

spatialCV_metrics %>%
  formattable(table,
            align=c("l","l"))
```

# 2021 Forecast
## Demande-side Change
### Population change

```{r pop}
countyPopulation_2021 <- 
  dat %>%
       st_set_geometry(NULL) %>%
       group_by(NAME) %>%
       summarize(county_population_2011 = round(sum(pop_2010))) %>%
  left_join(
    data.frame(
   NAME = 
     c("Fayette","Washington","Westmoreland","Allegheny","Beaver","Armstrong","Butler"),
   county_projection_2021 = 
     c(129274, 206865, 58002, 1216045, 163929, 64735, 187853))
  )

# Data source: 2019 estimate from upitt https://www.ucsur.pitt.edu/perspectives.php?b=20221115818670
```
```{r pop proj, include=T}
countyPopulation_2021 %>%
  gather(Variable,Value, -NAME) %>%
  ggplot(aes(reorder(NAME,-Value),Value)) +
  geom_bar(aes(fill=Variable), stat = "identity", position=position_dodge()) +
  scale_fill_manual(values = palette2,
                    labels=c("2011","2021"),
                    name="Population") +
  labs(title="Population Change by county: 2011 - 2021",
       x="County", y="Population") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  plotTheme()
```

### Land cover change
To picture 2021 land use scenario, we will project the change base on 2001-2011 pattern.

```{r reclasslc11}
developed11 <- lc_2011 == 21 | lc_2011 == 22 | lc_2011 == 23 | lc_2011 == 24
forest11 <- lc_2011 == 41 | lc_2011 == 42 | lc_2011 == 43 
farm11 <- lc_2011 == 81 | lc_2011 == 82 
wetlands11 <- lc_2011 == 90 | lc_2011 == 95 
otherUndeveloped11 <- lc_2011 == 52 | lc_2011 == 71 | lc_2011 == 31 
water11 <- lc_2011 == 11

names(developed11) <- "developed11"
names(forest11) <- "forest11"
names(farm11) <- "farm11"
names(wetlands11) <- "wetlands11"
names(otherUndeveloped11) <- "otherUndeveloped11"
names(water11) <- "water11"
```
```{r }
layer_list_11 <- list(developed11, wetlands11, forest11, farm11, otherUndeveloped11, water11)
names(layer_list_11) <- c("developed_2011", "wetlands_2011", "forest_2011", 
                       "farm_2011", "otherUndeveloped_2011", "water_2011")

fish_extract <- function(fishnet, layers) {
  extract_list <- exactextractr::exact_extract(layers, fishnet, fun = "mode")
  return(extract_list)
}


lc_2011_extracts <- lapply(layer_list_11, fish_extract, fishnet = MSA_fishnet)
lc_2011_extracts <- data.frame(do.call(cbind, args = lc_2011_extracts))
MSA_fishnet11 <- cbind(data.frame(MSA_fishnet), lc_2011_extracts)

if ( any("pop_2000.x" == names(MSA_fishnet11))) {
  MSA_fishnet11 <- 
    MSA_fishnet11 %>%
    mutate(pop_2000 = pop_2000.x,
         pop_2010 = pop_2010.x,
         pop_Change = pop_Change.x)
} 

lc_2011_lcvar <-
  MSA_fishnet11 %>%
  gather(var,value,developed_2011:water_2011) %>%
  dplyr::select(var,value,geometry) %>%
  st_as_sf()

lc_2011_lcvar$var <- as.factor(str_sub(lc_2011_lcvar$var,1,-6))
```
```{r lc type plot2, include=T}
# Plot
ggplot()+
  geom_sf(data=lc_2011_lcvar, aes(fill=as.factor(value)),color='transparent')+
  facet_wrap(~var)+
  scale_fill_manual(values=palette2, labels=c("Other","Land Cover"),name ="")+
  labs(title = "Land cover types, 2011",
         subtitle = "As fishnet centroids") +
  mapTheme()
```

```{r}
MSA_fishnet11 <-
  MSA_fishnet11 %>%
  rownames_to_column("fishnetID") %>%
  mutate(fishnetID = as.numeric(fishnetID)) 

MSA_fishnet11 <- left_join(MSA_fishnet11, MSA_fishnet01, by="fishnetID") %>%
  st_as_sf()

MSA_fishnet11 <-
  MSA_fishnet11 %>%
  dplyr::select(-developed_2001, -wetlands_2001, -forest_2001, -otherUndeveloped_2001, -farm_2001, -water_2001)


MSA_fishnet11$lagDevelopment <-
  nn_function(xyC(MSA_fishnet11),
              xyC(filter(MSA_fishnet11,developed_2011==1)),
              2)

colnames(MSA_fishnet11)[which (names(MSA_fishnet11)=="developed_0111")] <-"lc_change"

MSA_fishnet11 <- MSA_fishnet11 %>%
  mutate(developed11 = ifelse(lc_change==1 & developed_2011 == 1, 0, developed_2011)) #no use

dat11 <-
  MSA_fishnet11 %>%
  dplyr::select(lc_change, developed_2011,wetlands_2011,forest_2011,farm_2011,otherUndeveloped_2011,water_2011,pop_2000,pop_2010, pop_Change,
                distance_highways, lagDevelopment)  %>%
  st_join(Pittsburgh_MSA) %>%
  filter(water_2011 == 0)

# Force variable names to be same with model
dat11 <-
  dat11 %>%
  dplyr::mutate(developed_2001 = developed_2011,
                wetlands_2001 = wetlands_2011,
                forest_2001 = forest_2011,
                farm_2001 = farm_2011,
                otherUndeveloped_2001 = otherUndeveloped_2011) %>%
  dplyr::select(-developed_2011, -wetlands_2011, -forest_2011, -farm_2011, -otherUndeveloped_2011)


```

```{r demand scenario, include=T}
dat_infill <-
  dat11 %>%
  #calculate population change
    left_join(countyPopulation_2021) %>%
    mutate(proportion_of_county_pop = pop_2010 / county_population_2011,
           pop_2021.infill = proportion_of_county_pop * county_projection_2021,
           pop_Change = round(pop_2021.infill - pop_2010),2) %>%
    dplyr::select(-county_projection_2021, -county_population_2011, 
                  -proportion_of_county_pop, -pop_2021.infill) %>%
  #predict for 2021; only changed pop and lag
    mutate(predict_2021.infill = predict(Model6,. , type="response"))

dat_infill %>%
  ggplot() +  
  geom_point(aes(x=xyC(dat_infill)[,1], y=xyC(dat_infill)[,2], colour = factor(ntile(predict_2021.infill,5)))) +
  scale_colour_manual(values = palette5,
                    labels=substr(quintileBreaks(dat_infill, "predict_2021.infill"),1,8),
                    name="Quintile\nBreaks") +
  geom_sf(data=Pittsburgh_MSA, fill=NA, colour="black", size=1.5) +
  labs(title= "Development Demand in 2021: Predicted Probabilities") +
  mapTheme()
```

## Impact

```{r lc1101 plot, include=T}
ggplot() +
  geom_sf(data=Pittsburgh_MSA_boundary) +
  geom_raster(data = rbind(rast(lc_2001) %>% mutate(label = "2001"),
                           rast(lc_2011) %>% mutate(label = "2011")) %>% 
              na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(value))) +
  facet_wrap(~label) +
  scale_fill_viridis(discrete=TRUE, direction=-1,name ="") +
  labs(title = "Land Cover, 2001 & 2011") +
  mapTheme() + theme(legend.position = "none")
```

```{r sensitive land, include=T}
dat0111 <- st_join(MSA_fishnet01, MSA_fishnet11, by="geometry")

dat2 <-
  dat0111 %>%
   mutate(sensitive_lost11 = ifelse(forest_2001 == 1 & forest_2011 == 0 |
                                    wetlands_2001 == 1 & wetlands_2011 == 0,1,0)) # this includes lands that are not developed at all (farm)
                      
ggplot() +
  geom_point(data=dat2, aes(x=xyC(dat2)[,1], y=xyC(dat2)[,2], colour=as.factor(sensitive_lost11))) +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","Sensitive Lost"),
                      name = "") +
  labs(title = "Sensitive lands lost: 2001 - 2011",
       subtitle = "As fishnet centroids") +
  mapTheme()
```

```{r land fragment, include=T}
sensitiveRegions <- 
  raster::clump(wetlands11 + forest11) %>%
  rasterToPolygons() %>%
  st_as_sf() %>%
  group_by(clumps) %>% summarize() %>%
    mutate(Acres = as.numeric(st_area(.) * 0.0000229568)) %>%
    filter(Acres > 3954)  %>%
  dplyr::select() %>%
  rasterize(.,emptyRaster) 
sensitiveRegions[sensitiveRegions > 0] <- 1  
names(sensitiveRegions) <- "sensitiveRegions"

dat2 <-
  aggregateRaster(c(sensitiveRegions), dat2) %>%
  dplyr::select(sensitiveRegions) %>%
  st_set_geometry(NULL) %>%
  bind_cols(.,dat2) %>%
  st_sf()

ggplot() +
  geom_point(data=dat2, aes(x=xyC(dat2)[,1], y=xyC(dat2)[,2], colour=as.factor(sensitiveRegions))) +
  scale_colour_manual(values = palette2,
                      labels=c("Other","Sensitive Regions"),
                      name="") +
  labs(title = "Sensitive regions",
       subtitle = "Continous areas of either wetlands or forests\ngreater than 1 acre") +
  mapTheme()
```

```{r dat2 process}

```


```{r demand county}
county_specific_metrics <- 
  dat2 %>%
  #predict development demand from our model
  mutate(Development_Demand = predict(Model6, dat2, type="response")) %>% #change name of variable!
  #get a count count of grid cells by county which we can use to calculate rates below
  left_join(st_set_geometry(dat, NULL) %>% group_by(NAME) %>% summarize(count = n()), by = character()) %>%
  #calculate summary statistics by county
  group_by(NAME) %>%
  summarize(Total_Farmland = sum(farm_2011) / max(count),
            Total_Forest = sum(forest_2011) / max(count),
            Total_Wetlands = sum(wetlands_2011) / max(count),
            Total_Undeveloped = sum(otherUndeveloped_2011) / max(count),
            Sensitive_Land_Lost = sum(sensitive_lost11) / max(count),
            Sensitive_Regions = sum(sensitiveRegions) / max(count),
            Mean_Development_Demand = mean(Development_Demand)) %>%
  #get population data by county
  left_join(countyPopulation_2021 %>% 
            mutate(Population_Change = county_projection_2021 - county_population_2011,
                   Population_Change_Rate = Population_Change / county_projection_2021) %>%
            dplyr::select(NAME,Population_Change_Rate))

county_specific_metrics <- 
  dat_infill %>%
  left_join(st_set_geometry(dat, NULL) %>% group_by(NAME) %>% summarize(count = n()), by = character()) %>%
  #calculate summary statistics by county
  group_by(NAME) %>%
  summarize(Total_Farmland = sum(farm_2011) / max(count),
            Total_Forest = sum(forest_2011) / max(count),
            Total_Wetlands = sum(wetlands_2011) / max(count),
            Total_Undeveloped = sum(otherUndeveloped_2011) / max(count),
            Sensitive_Land_Lost = sum(sensitive_lost11) / max(count),
            Sensitive_Regions = sum(sensitiveRegions) / max(count),
            Mean_Development_Demand = mean(Development_Demand)) %>%
  #get population data by county
  left_join(countyPopulation_2021 %>% 
            mutate(Population_Change = county_projection_2021 - county_population_2011,
                   Population_Change_Rate = Population_Change / county_projection_2021) %>%
            dplyr::select(NAME,Population_Change_Rate))
```

```{r demand county plot, include=T}
county_specific_metrics %>%
  gather(Variable, Value, -NAME, -geometry) %>%
  mutate(Variable = factor(Variable, levels=c("Population_Change_Rate","Mean_Development_Demand",
                                              "Total_Farmland","Total_Undeveloped","Total_Forest",
                                              "Total_Wetlands","Sensitive_Land_Lost","Sensitive_Regions",
                                              ordered = TRUE))) %>%
  mutate(Planning_Designation = case_when(
    Variable == "Population_Change_Rate" | Variable == "Mean_Development_Demand" ~ "Demand-Side",
    Variable == "Total_Farmland" | Variable == "Total_Undeveloped"               ~ "Suitable",
    TRUE                                                                         ~ "Not Suitable")) %>%
  ggplot(aes(x=Variable, y=Value, fill=Planning_Designation)) +
    geom_bar(stat="identity", position=position_dodge(), colour="black") +
    facet_wrap(~NAME, ncol=5) +
    coord_flip() +
    scale_y_continuous(breaks = seq(.25, 1, by = .25)) +
    geom_vline(xintercept = 2.5) + geom_vline(xintercept = 4.5) +
    scale_fill_manual(values=c("black","red","darkgreen")) +
    labs(title= "County Specific Allocation Metrics", subtitle= "As rates", x="Indicator", y="Rate") +
    plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="bottom")
```

## Supply-side Change


# Allocation

```{r allocation, include=T}
Allegheny <-
  dat2 %>%
    mutate(Development_Demand = predict(Model6, dat2, type="response")) %>%
    left_join(st_set_geometry(dat, NULL) %>% group_by(NAME) %>% summarize(count = n()), by = character()) %>%
    #calculate summary statistics by county
    filter(NAME == "Allegheny") 

Allegheny_landUse <- rbind(
  filter(Allegheny, forest_2011 == 1 | wetlands_2011 == 1 ) %>%
  dplyr::select() %>% mutate(Land_Use = "Not Suitable"),
  filter(Allegheny, developed_2011 == 1) %>%
  dplyr::select() %>% mutate(Land_Use = "Developed"))

grid.arrange(
ggplot() +
  geom_sf(data=Allegheny, aes(fill=factor(ntile(Development_Demand,5))), colour=NA) +
  geom_point(data=Allegheny_landUse, aes(x=xyC(Allegheny_landUse)[,1], 
                                        y=xyC(Allegheny_landUse)[,2], colour=Land_Use),
                                        shape = 15, size = 2) +
  geom_sf(data=st_intersection(PittsburghHighways,filter(Pittsburgh_MSA, NAME=="Allegheny")), size=2) +
  scale_fill_manual(values = palette5, name="Development\nDemand",
                    labels=substr(quintileBreaks(Allegheny,"Development_Demand"),1,5)) +
  scale_colour_manual(values = c("black","red")) + 
  labs(title = "Development Potential, 2021: Allegheny") + mapTheme() +
  guides(fill = guide_legend(order = 1), colour = guide_legend(order = 2)),

ggplot() +
  geom_sf(data=Allegheny, aes(fill=factor(ntile(pop_Change,5))), colour=NA) +
  geom_point(data=Allegheny_landUse, aes(x=xyC(Allegheny_landUse)[,1], 
                                        y=xyC(Allegheny_landUse)[,2], colour=Land_Use),
                                        shape = 15, size = 2) +
  geom_sf(data=st_intersection(PittsburghHighways,filter(Pittsburgh_MSA, NAME=="Allegheny")), size=2) +
  scale_fill_manual(values = palette5, name="Population\nChange",
                    labels=substr(quintileBreaks(Allegheny,"pop_Change"),1,5)) +
  scale_colour_manual(values = c("black","red")) + 
  labs(title = "Projected Population, 2021: Allegheny") + mapTheme() +
  guides(fill = guide_legend(order = 1), colour = guide_legend(order = 2)), ncol=2)
```


# Appendices {.tabset .tabset-fade}
## Data Dictionary
### Land cover change data
<p>The table below shows descriptions of each categorical land cover type in the land cover data, which corresponds to the numeric values.</p>
<table class="table" style="font-size: 10px; margin-left: auto; margin-right: auto;">
<thead>
<tr>
<th style="text-align:left;">
Definition
</th>
<th style="text-align:left;">
Value
</th>
</tr>
</thead>
<tbody>
<tr>
<td style="text-align:left;width: 50em; ">
Open Water - All areas of open water, generally with less than 25% cover or vegetation or soil.
</td>
<td style="text-align:left;width: 10em; ">
11
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Perennial Ice/Snow - All areas characterized by a perennial cover of ice and/or snow, generally greater than 25% of total cover.
</td>
<td style="text-align:left;width: 10em; ">
12
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Developed, Open Space - Includes areas with a mixture of some constructed materials, but mostly vegetation in the form of lawn grasses. Impervious surfaces account for less than 20 percent of total cover. These areas most commonly include large-lot single-family housing units, parks, golf courses, and vegetation planted in developed settings for recreation, erosion control, or aesthetic purposes.
</td>
<td style="text-align:left;width: 10em; ">
21
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Developed, Low Intensity -Includes areas with a mixture of constructed materials and vegetation. Impervious surfaces account for 20-49 percent of total cover. These areas most commonly include single-family housing units.
</td>
<td style="text-align:left;width: 10em; ">
22
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Developed, Medium Intensity - Includes areas with a mixture of constructed materials and vegetation. Impervious surfaces account for 50-79 percent of the total cover. These areas most commonly include single-family housing units.
</td>
<td style="text-align:left;width: 10em; ">
23
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Developed, High Intensity - Includes highly developed areas where people reside or work in high numbers. Examples include apartment complexes, row houses and commercial/industrial. Impervious surfaces account for 80 to 100 percent of the total cover.
</td>
<td style="text-align:left;width: 10em; ">
24
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Barren Land (Rock/Sand/Clay) - Barren areas of bedrock, desert pavement, scarps, talus, slides, volcanic material, glacial debris, sand dunes, strip mines, gravel pits and other accumulations of earthen material. Generally, vegetation accounts for less than 15% of total cover.
</td>
<td style="text-align:left;width: 10em; ">
31
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Deciduous Forest - Areas dominated by trees generally greater than 5 meters tall, and greater than 20% of total vegetation cover. More than 75 percent of the tree species shed foliage simultaneously in response to seasonal change.
</td>
<td style="text-align:left;width: 10em; ">
41
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Evergreen Forest - Areas dominated by trees generally greater than 5 meters tall, and greater than 20% of total vegetation cover. More than 75 percent of the tree species maintain their leaves all year. Canopy is never without green foliage.
</td>
<td style="text-align:left;width: 10em; ">
42
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Mixed Forest - Areas dominated by trees generally greater than 5 meters tall, and greater than 20% of total vegetation cover. Neither deciduous nor evergreen species are greater than 75 percent of total tree cover.
</td>
<td style="text-align:left;width: 10em; ">
43
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Dwarf Scrub - Alaska only areas dominated by shrubs less than 20 centimeters tall with shrub canopy typically greater than 20% of total vegetation. This type is often co-associated with grasses, sedges, herbs, and non-vascular vegetation.
</td>
<td style="text-align:left;width: 10em; ">
51
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Shrub/Scrub - Areas dominated by shrubs; less than 5 meters tall with shrub canopy typically greater than 20% of total vegetation. This class includes true shrubs, young trees in an early successional stage or trees stunted from environmental conditions.
</td>
<td style="text-align:left;width: 10em; ">
52
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Grassland/Herbaceous - Areas dominated by grammanoid or herbaceous vegetation, generally greater than 80% of total vegetation. These areas are not subject to intensive management such as tilling, but can be utilized for grazing.
</td>
<td style="text-align:left;width: 10em; ">
71
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Sedge/Herbaceous - Alaska only areas dominated by sedges and forbs, generally greater than 80% of total vegetation. This type can occur with significant other grasses or other grass like plants, and includes sedge tundra, and sedge tussock tundra.
</td>
<td style="text-align:left;width: 10em; ">
72
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Lichens - Alaska only areas dominated by fruticose or foliose lichens generally greater than 80% of total vegetation.
</td>
<td style="text-align:left;width: 10em; ">
73
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Moss - Alaska only areas dominated by mosses, generally greater than 80% of total vegetation.
</td>
<td style="text-align:left;width: 10em; ">
74
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Pasture/Hay - Areas of grasses, legumes, or grass-legume mixtures planted for livestock grazing or the production of seed or hay crops, typically on a perennial cycle. Pasture/hay vegetation accounts for greater than 20 percent of total vegetation.
</td>
<td style="text-align:left;width: 10em; ">
81
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Cultivated Crops - Areas used for the production of annual crops, such as corn, soybeans, vegetables, tobacco, and cotton, and also perennial woody crops such as orchards and vineyards. Crop vegetation accounts for greater than 20 percent of total vegetation. This class also includes all land being actively tilled.
</td>
<td style="text-align:left;width: 10em; ">
82
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Woody Wetlands - Areas where forest or shrub land vegetation accounts for greater than 20 percent of vegetative cover and the soil or substrate is periodically saturated with or covered with water.
</td>
<td style="text-align:left;width: 10em; ">
90
</td>
</tr>
<tr>
<td style="text-align:left;width: 50em; ">
Emergent Herbaceous Wetlands - Areas where perennial herbaceous vegetation accounts for greater than 80 percent of vegetative cover and the soil or substrate is periodically saturated with or covered with water.
</td>
<td style="text-align:left;width: 10em; ">
95
</td>
</tr>
</tbody>
<tfoot>
<tr>
<td style="padding: 0; border: 0;" colspan="100%">
<span style="font-style: italic;">Source</span>
</td>
</tr>
<tr>
<td style="padding: 0; border: 0;" colspan="100%">
<sup></sup> <a href="https://www.mrlc.gov/data/nlcd-2001-2011-land-cover-change-conus" class="uri">https://www.mrlc.gov/data/nlcd-2001-2011-land-cover-change-conus</a>
</td>
</tr>
</tfoot>
</table>

### Land Cover Classification
<p> The table below shows how we reclassify land cover data into more useful categories.</p>
<table class="table" style="width: auto !important; margin-left: auto; margin-right: auto;">
<thead>
<tr>
<th style="text-align:left;">
Old_Classification
</th>
<th style="text-align:left;">
New_Classification
</th>
</tr>
</thead>
<tbody>
<tr>
<td style="text-align:left;">
Open Space as well as Low, Medium and High Intensity Development
</td>
<td style="text-align:left;">
Developed
</td>
</tr>
<tr>
<td style="text-align:left;">
Deciduous, Evergreen and Mixed Forest
</td>
<td style="text-align:left;">
Forest
</td>
</tr>
<tr>
<td style="text-align:left;">
Pasture/Hay and Cultivated Crops
</td>
<td style="text-align:left;">
Farm
</td>
</tr>
<tr>
<td style="text-align:left;">
Woody and Emergent Herbaceous Wetlands
</td>
<td style="text-align:left;">
Woodlands
</td>
</tr>
<tr>
<td style="text-align:left;">
Barren Land, Dwarf Scrub and Grassland/Herbaceous
</td>
<td style="text-align:left;">
Other Undeveloped
</td>
</tr>
<tr>
<td style="text-align:left;">
Water
</td>
<td style="text-align:left;">
Water
</td>
</tr>
</tbody>
</table>


## Code
### Preparation
```{r prep, echo=T, include=T, eval=F}
library(tidyverse)
library(sf)
library(raster)
library(knitr)
library(formattable)
library(tidycensus)
library(tigris)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
library(FNN)
library(QuantPsyc)
library(caret)
library(yardstick)
library(pscl)
library(plotROC) 
library(ggrepel)
library(pROC)
library(grid)
library(gridExtra)
library(viridis)
library(exactextractr)
library(stringr)
library(censusapi)
library(ROCR)

mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2)
  )
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 14)
  )
}

palette2 <- c("#41b6c4","#253494")
palette4 <- c("#a1dab4","#41b6c4","#2c7fb8","#253494")
palette5 <- c("#ffffcc","#a1dab4","#41b6c4","#2c7fb8","#253494")
palette10 <- c("#f7fcf0","#e0f3db","#ccebc5","#a8ddb5","#7bccc4",
               "#4eb3d3","#2b8cbe","#0868ac","#084081","#f7fcf0")
```
```{r function prep, echo=T, include=T, eval=F}
quintileBreaks <- function(df,variable) {
    as.character(quantile(df[[variable]],
                          c(.01,.2,.4,.6,.8),na.rm=T))
}

#This function can be used to convert a polygon sf to centroids xy coords.
xyC <- function(aPolygonSF) {
  as.data.frame(
    cbind(x=st_coordinates(st_centroid(aPolygonSF))[,1],
          y=st_coordinates(st_centroid(aPolygonSF))[,2]))
} 

#this function convert a raster to a data frame so it can be plotted in ggplot
rast <- function(inRaster) {
  data.frame(
    xyFromCell(inRaster, 1:ncell(inRaster)), 
    value = getValues(inRaster)) }

nn_function <- function(measureFrom,measureTo,k) {
  #convert the sf layers to matrices
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
    output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}
```
### Land Use Change Data and Create Fishnet
```{r load and manipulate lc data, echo=T, include=T, eval=F}
# Xinyi's
# Pittsburgh_MSA <-
#   st_read("Pittsburgh_MSA/Pittsburgh_MSA.shp") %>%
#   st_transform(102741) #102003
# 
# Pittsburgh_MSA_boundary <- Pittsburgh_MSA %>%
#   group_by(STATEFP) %>%
#   summarise() %>%
#   st_transform(102741)

# Gillian's
Pittsburgh_MSA <-
  st_read("Pittsburgh_MSA/Pittsburgh_MSA.shp") %>%
  st_transform("ESRI:102741") #102003

Pittsburgh_MSA_boundary <- Pittsburgh_MSA %>%
  group_by(STATEFP) %>%
  summarise() %>%
  st_transform("ESRI:102741")

MSA_fishnet <- 
  st_make_grid(Pittsburgh_MSA_boundary, 3000) %>%
  st_sf()

MSA_fishnet <- 
  MSA_fishnet[Pittsburgh_MSA_boundary,]

ggplot()+
  labs(title="Pittsburgh MSA Boundary and Counties",
       subtitle="Illustration of Fishnet, 3000 foot resolution")+
  geom_sf(data=Pittsburgh_MSA, aes(fill=as.factor(NAME)))+
  geom_sf(data=MSA_fishnet, fill=NA, color="gray", alpha=0.5) +
  geom_sf(data=Pittsburgh_MSA, fill=NA, size=1.5, aes(colour=as.factor(NAME)))+
  scale_color_viridis(discrete = T,
                     name ="County")+
  scale_fill_viridis(discrete = T,
                     name ="County")+
  mapTheme()

lc_2001 <- raster("msa_2001lc.tif")
lc_2001 <-projectRaster(lc_2001,crs=crs(Pittsburgh_MSA))

lc_2011 <- raster ("msa_2011lc.tif")
lc_2011 <- projectRaster(lc_2011,crs=crs(Pittsburgh_MSA) )

# aggregate
lc_2001 <- raster::aggregate(lc_2001, fact = 30, fun = modal)
lc_2011 <- raster::aggregate(lc_2011, fact = 30, fun = modal)

# calculate
lc_change <- (lc_2001 != lc_2011) * lc_2011

reclassMatrix <- 
  matrix(c(
    0,12,0,
    12,24,1,
    24,Inf,0),
    ncol=3, byrow=T)
lc_change_reclass <- reclassify(lc_change, reclassMatrix)
lc_change_reclass[lc_change_reclass < 1] <- 0
names(lc_change_reclass) <- "lc_change_reclass"

lc_change_extract <- data.frame(exact_extract(lc_change_reclass, MSA_fishnet, fun = "mode"))
names(lc_change_extract) <- "developed_2011"

MSA_fishnet <- cbind(data.frame(MSA_fishnet),lc_change_extract)

# Xinyi's
# MSA_fishnet <- 
#   MSA_fishnet %>%
#   st_as_sf(.) %>%
#   st_transform(102741)

MSA_fishnet <- 
  MSA_fishnet %>%
  st_as_sf(.) %>%
  st_transform("ESRI:102741")

# Plot
ggplot() +
   geom_sf(data=MSA_fishnet,aes(fill=as.factor(developed_2011)),color='transparent')+
   scale_fill_viridis(discrete=TRUE, name ="Land Cover\nChange",labels=c("No Change","New Development")) +
  labs(title="Pittsburgh MSA Development Land Use Change (2001 - 2011)") +
  mapTheme()

ggplot() +
  geom_sf(data=Pittsburgh_MSA_boundary) +
  geom_raster(data=rast(lc_2001) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(round(value,0)))) +
  scale_fill_viridis(discrete=TRUE,direction=-1, name ="") +
  labs(title = "Pittsburgh MSA Land Cover, 2001") +
  mapTheme()

developed <- lc_2001 == 21 | lc_2001 == 22 | lc_2001 == 23 | lc_2001 == 24
forest <- lc_2001 == 41 | lc_2001 == 42 | lc_2001 == 43 
farm <- lc_2001 == 81 | lc_2001 == 82 
wetlands <- lc_2001 == 90 | lc_2001 == 95 
otherUndeveloped <- lc_2001 == 52 | lc_2001 == 71 | lc_2001 == 31 
water <- lc_2001 == 11
names(developed) <- "developed"
names(forest) <- "forest"
names(farm) <- "farm"
names(wetlands) <- "wetlands"
names(otherUndeveloped) <- "otherUndeveloped"
names(water) <- "water"

layer_list <- list(developed, wetlands, forest, farm, otherUndeveloped, water)
names(layer_list) <- c("developed_2001", "wetlands_2001", "forest_2001", 
                       "farm_2001", "otherUndeveloped_2001", "water_2001")

fish_extract <- function(fishnet, layers) {
  
  extract_list <- exact_extract(layers, fishnet, fun = "mode")
  
  return(extract_list)
  
}


lc_2001_extracts <- lapply(layer_list, fish_extract, fishnet = MSA_fishnet)
lc_2001_extracts <- data.frame(do.call(cbind, args = lc_2001_extracts))
MSA_fishnet <- cbind(data.frame(MSA_fishnet), lc_2001_extracts)

# Xinyi's
# MSA_fishnet <- 
#   MSA_fishnet %>%
#   st_as_sf(.) %>%
#   st_transform(102741)

MSA_fishnet <- 
  MSA_fishnet %>%
  st_as_sf(.) %>%
  st_transform("ESRI:102741")

lc_2001_lcvar <-
  MSA_fishnet %>%
  gather(var,value,developed_2001:water_2001) %>%
  dplyr::select(var,value,geometry) 

lc_2001_lcvar$var <- as.factor(str_sub(lc_2001_lcvar$var,1,-6))

# Plot
ggplot()+
  geom_sf(data=lc_2001_lcvar,aes(fill=as.factor(value)),color='transparent')+
  facet_wrap(~var)+
  scale_fill_viridis(discrete=TRUE, labels=c("Other","Land Cover"),name ="")+
    labs(title = "Land cover types, 2001",
         subtitle = "As fishnet centroids") +
   mapTheme()
```
### Predicting Features Data
```{r feature data, echo=T, include=T, eval=F}
# Population
MSAPop10 <-
  get_decennial(geography = "tract",variables = "P001001",year=2010,
                state=42,county=c("Allegheny County","Armstrong County","Beaver County","Butler County","Fayette County",
                                  "Washington County","Westmoreland County"),geometry=TRUE) %>%
  rename(pop_2010 = value) %>%
  st_transform(st_crs(MSA_fishnet))

Sys.setenv(CENSUS_KEY = "2c3e9f9d2d65abab5f7b81fe418054415d363a43")
MSAPop00 <-getCensus(name = "dec/sf1", 
                  vintage = 2000, 
                  vars = c("NAME","P001001"),
                  region = "tract:*", 
                  regionin = "state:42+county:003,005,007,019,051,125,129")
MSAPop00$label <- paste0(MSAPop00$county,"0",MSAPop00$tract)
for (i in 1:721){
  if(nchar(MSAPop00[i,]$label)!=10){
    MSAPop00[i,]$label <-paste0(MSAPop00[i,]$label,"00")
  }
  
}

# Xinyi's 
# MSAtracts00 <- st_read("MSAtracts/MSAtracts.shp") %>%
#  st_transform(102741)

# Gillian's
MSAtracts00 <- st_read("MSAtracts/MSAtracts.shp") %>%
  st_transform("ESRI:102741")

MSAtracts00$tract <-substr(MSAtracts00$GISJOIN2 ,4,13)
MSAtracts00$tract <- as.character(MSAtracts00$tract)

MSAPop00 <- left_join(MSAtracts00, MSAPop00,by=c("tract"="label"))
MSAPop00 <- MSAPop00 %>%
  dplyr::select(GISJOIN2,NAME,P001001,geometry) %>%
  rename(GEOID = GISJOIN2,pop_2000 = P001001)

# Plot
grid.arrange(
ggplot() +
  geom_sf(data = MSAPop00, aes(fill=factor(ntile(pop_2000,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=quintileBreaks(MSAPop00,"pop_2000"),
                   name="Quintile\nBreaks") +
  labs(title="Population, Pittsburgh MSA: 2000") +
  mapTheme(),

ggplot() +
  geom_sf(data = MSAPop10, aes(fill=factor(ntile(pop_2010,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=quintileBreaks(MSAPop10,"pop_2010"),
                   name="Quintile\nBreaks") +
  labs(title="Population, Pittsburgh MSA: 2010") +
  mapTheme(), ncol=2)

# To fishnet
MSA_fishnet <-
  MSA_fishnet %>%
  rownames_to_column("fishnetID") %>%
  mutate(fishnetID = as.numeric(fishnetID)) 

fishnetPop00 <-
  st_interpolate_aw(MSAPop00["pop_2000"],MSA_fishnet,extensive=TRUE) %>%
  as.data.frame(.) %>%
  left_join(MSA_fishnet,.,by="geometry") %>%
  mutate(pop_2000=replace_na(pop_2000,0)) %>%
  dplyr::select(fishnetID,pop_2000)

fishnetPop10 <-
  st_interpolate_aw(MSAPop10["pop_2010"],MSA_fishnet,extensive=TRUE) %>%
  as.data.frame(.) %>%
  left_join(MSA_fishnet,.,by="geometry") %>%
  mutate(pop_2010=replace_na(pop_2010,0)) %>%
  dplyr::select(pop_2010,fishnetID)

fishnetPop <- 
  cbind(fishnetPop00,fishnetPop10) %>%
  dplyr::select(pop_2000,pop_2010,fishnetID) %>%
  mutate(pop_Change = pop_2010 - pop_2000) %>%
  st_drop_geometry(.)


MSA_fishnet <- left_join(MSA_fishnet,data.frame(fishnetPop),by="fishnetID")

# Highway
PittsburghHighways <-
  st_read("MSAHighways/MSAHighways.shp") %>%
  st_transform(st_crs(Pittsburgh_MSA)) %>%
  st_intersection(Pittsburgh_MSA_boundary)

PittsburghHighways <- 
  PittsburghHighways %>%
  dplyr::select(CTY_CODE, DISTRICT_N,MAINT_FUNC,geometry)

# Plot
ggplot() +
  geom_sf(data=MSA_fishnet,aes(fill=as.factor(developed_2011)),color='transparent') +
  geom_sf(data=PittsburghHighways,color='white') +
  scale_fill_viridis(discrete=TRUE, name ="Land Cover\nChange",labels=c("No Change","New Development")) +
  labs(title = "Pittsburgh MSA New Development and Highways") +
  mapTheme()

# Distance to highway

emptyRaster <- lc_change
emptyRaster[] <- NA

highway_raster <- 
  as(PittsburghHighways,'Spatial') %>%
  rasterize(.,emptyRaster)

# writeRaster(highway_raster,"highway_raster.tif")

highway_raster_distance <- distance(highway_raster)
names(highway_raster_distance) <- "distance_highways"

highwayPoints <-
  rasterToPoints(highway_raster_distance) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(MSA_fishnet))

highwayPoints_fishnet <- 
  aggregate(highwayPoints, MSA_fishnet, mean) %>%
  mutate(distance_highways = ifelse(is.na(distance_highways),0,distance_highways)) %>%
  rownames_to_column("fishnetID") %>%
  mutate(fishnetID = as.numeric(fishnetID)) 
  

highwayPoints_fishnet <-highwayPoints_fishnet %>%
  st_drop_geometry(.)

MSA_fishnet <- left_join(MSA_fishnet,highwayPoints_fishnet,by="fishnetID")

# Plot dist to highway
ggplot() +
  geom_sf(data=Pittsburgh_MSA_boundary) +
  geom_point(data=MSA_fishnet , aes(x=xyC(MSA_fishnet )[,1], 
                                             y=xyC(MSA_fishnet)[,2], 
                 colour=factor(ntile(distance_highways,5))),size=1.5) +
  scale_colour_manual(values = palette5,
                      labels=substr(quintileBreaks(MSA_fishnet ,"distance_highways"),1,8),
                      name="Quintile\nBreaks") +
  geom_sf(data=PittsburghHighways, colour = "red") +
  labs(title = "Distance to Highways",
       subtitle = "As fishnet centroids; Highways visualized in red") +
  mapTheme()

# Spatial lag of devlopment
MSA_fishnet$lagDevelopment <-
  nn_function(xyC(MSA_fishnet),
              xyC(filter(MSA_fishnet,developed_2001==1)),
              2)

# Plot nn functions
ggplot() +
  geom_sf(data=Pittsburgh_MSA_boundary) +
  geom_point(data=MSA_fishnet, 
             aes(x=xyC(MSA_fishnet)[,1], y=xyC(MSA_fishnet)[,2], 
                 colour=factor(ntile(lagDevelopment,5))), size=1.5) +
  scale_colour_manual(values = palette5,
                     labels=substr(quintileBreaks(MSA_fishnet,"lagDevelopment"),1,7),
                     name="Quintile\nBreaks") +
  labs(title = "Spatial lag to 2001 development",
       subtitle = "As fishnet centroids") +
  mapTheme()
```
```{r exploratory, echo=T, include=T, eval=F}
# Final dataset
colnames(MSA_fishnet)[which (names(MSA_fishnet)=="developed_2011")] <-"lc_change"

dat <-
  MSA_fishnet %>%
  dplyr::select(lc_change,developed_2001,wetlands_2001,forest_2001,farm_2001,otherUndeveloped_2001,water_2001,pop_2000,pop_2010,pop_Change,
                distance_highways, lagDevelopment)  %>%
  st_join(Pittsburgh_MSA) %>%
  mutate(developed11 = ifelse(lc_change==1 & developed_2001 == 1,0,developed_2001)) %>%
  filter(water_2001 == 0)

dat$lc_change <- as.factor(dat$lc_change)

dat %>%
  dplyr::select(distance_highways,lagDevelopment,lc_change) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun = "mean") +
    facet_wrap(~Variable) +
    scale_fill_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New development as a function of the countinuous variables") +
    plotTheme() 

grid.arrange(
dat %>%
  dplyr::select(pop_2000,pop_2010,pop_Change,lc_change) %>%
  mutate(pop_Change=pop_Change) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun = "mean") +
    facet_wrap(~Variable) +
    scale_fill_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New development as a function of factor variables",
         subtitle = "Original Population Change") +
    plotTheme(),

dat %>%
  dplyr::select(pop_2000,pop_2010,pop_Change,lc_change) %>%
  mutate(pop_Change=pop_Change*10) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun = "mean") +
    facet_wrap(~Variable) +
    scale_fill_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New development as a function of factor variables",
         subtitle = "Multiply Population Change by 10") +
    plotTheme(),ncol=1)

table <-dat %>%
  dplyr::select(lc_change:otherUndeveloped_2001) %>%
  gather(Land_Cover_Type, Value, -lc_change, -geometry) %>%
   st_set_geometry(NULL) %>%
     group_by(lc_change, Land_Cover_Type) %>%
     summarize(n = sum(as.numeric(Value))) %>%
     ungroup() %>%
    mutate(Conversion_Rate = paste0(round(100 * n/sum(n), 2), "%")) %>%
    filter(lc_change == 1) %>%
  dplyr::select(Land_Cover_Type,Conversion_Rate) 
table <- data.frame(table)
table$Land_Cover_Type[1]<-"Developed"
table$Land_Cover_Type[2]<-"Farm"
table$Land_Cover_Type[3]<-"Forest"
table$Land_Cover_Type[4]<-"Other Undeveloped"
formattable(table,
            align=c("l","l"))
```